---
title: "Neural Networks"
subtitle: "Project 2"
author: "Elliott Chang, Cameron Adams"
title-slide-attributes:
  data-slide-number: none
format: revealjs
editor: visual
execute:
  echo: false
---

```{r}
#| label: load-packages
#| include: false

library(tidyverse)
```

```{r}
#| label: setup
#| include: false

# For better figure resolution
knitr::opts_chunk$set(
  fig.retina = 3, 
  dpi = 300, 
  fig.width = 6, 
  fig.asp = 0.618, 
  out.width = "70%"
  )
```

```{r}
#| label: load-data
#| include: false
```

## Introduction

We plan on illustrating the predictive power and flexibility of Neural Networks on data with complex relationships.

-   For the sake of this project, we decided to focus on classification

-   

## The data:

To illustrate the power and flexibility of neural networks, we created multiple datasets with complex relationships.

-   Each dataset contains two continuous predictor variables `x` and `y`, and one categorical response variable `label` with factors `a` and `b`.

-   Two coninuous predictors allow us to visualize the relationships between these predictors easily, giving us an intuitive understand of what the underlying relationship might be

    -   Despite the intuitive nature of the relationships, modeling them is quite difficult using tools that we've learned about so far in this class

## What is a Neural Network?
